---
title: "Visualization Rueda-Alaña 2024"
author: "phylobrain"
date: "2023-11-07"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(Seurat)
```

## Import Seurat Object

```{r import}
path_to_bs_e12p3_all = "BirthSeq_E12-P3_AllCells.rds"
bs_e12p3_all <- readRDS(path_to_bs_e12p3_all)
```

## Pre-proccessing

The Seurat object has 3 different assays: RNA, raw counts; SCT, normalized; integrated, corrected batch effect. 
If you want to replicate this steps, this is the code. Subtle variations are expected due to version differences:

```
DefaultAssay(mp2_2)<-"RNA"
bs_e12p3.list <- SplitObject(mp2_2, split.by="orig.ident")
bs_e12p3.list <- lapply(X = bs_e12p3.list, FUN = SCTransform, method = "glmGamPoi",  vars.to.regress = c("S.Score", "G2M.Score","percent.mt"))
features <- SelectIntegrationFeatures(object.list = bs_e12p3.list, nfeatures = 10000)
bs_e12p3.list <- PrepSCTIntegration(object.list = bs_e12p3.list, anchor.features = features)
bs_e12p3.list <- lapply(X = bs_e12p3.list, FUN = RunPCA, features = features)
anchors <- FindIntegrationAnchors(object.list = bs_e12p3.list, reduction = "rpca", dims = 1:50, normalization.method = "SCT", anchor.features = features)
mp2_2 <- IntegrateData(anchorset = anchors, normalization.method = "SCT", dims = 1:50)
```

# Visualization 

Apart from the QC parameters, the following annotation can be found: 
"CellType": General annotation used for Fig5C

```{r metadata, echo=FALSE}
DimPlot(bs_e12p3_all, group.by = "CellType")
```

Visualization of marker genes for each cluster 

```{r}
genes_fig5d<-c("Pdgfra","Aldoc","Ascl1","Dlx1","Neurod2","Reln")
for(i in 1:length(genes_fig5d)){print(FeaturePlot(bs_e12p3_all, features = genes_fig5d[i]))}
```

# Neuron subsetting

For further study of the dataset and annotation, the neuronal population were subsetted based on previous annotation and discarded low quality clusters.

The data was again preprocessed with the recalculation of QC features.

To import the already proccessed object:

# Import 

```{r}
path_to_bs_e12p3_neurons = "BirthSeq_E12-P3_Neurons.rds"
bs_e12p3_neurons <- readRDS(path_to_bs_e12p3_neurons)
```

# Visualization 

Apart from the QC parameters, the following annotation can be found: 
"CellType": General annotation used for Fig5C
"Named_Subclustering": Subclustering annotation used for Fig5E

```{r metadata, echo=FALSE}
DimPlot(bs_e12p3_neurons, group.by = "CellType")
```

```{r metadata, echo=FALSE}
DimPlot(bs_e12p3_neurons, group.by = "Named_Subclustering")
```

Visualization of marker genes for each cluster 

```{r}
genes_fig5g<-c("Cxcl12","Trp73","Lhx5","Snhg11","Reln")
for(i in 1:length(genes_fig5g)){print(FeaturePlot(bs_e12p3_neurons, features = genes_fig5g[i]))}
```
