---
title: "GSI Correlation Function"
output: html_notebook
---

```{r}
library(dplyr)
library(Seurat)
library(SeuratObject)
library(patchwork)
library(Matrix)
library(glmGamPoi)
library(sctransform)
library(limma)
library(clustree)
library(pvclust)
library(Hmisc)
library(corrplot)
library(ggcorrplot)
library(stringr)
library(biomaRt)
source("/media/phylobrain/Windows/Users/rodri/Documents/R/win-library/4.1/corrplot/R/corrtry.R")
source("/home/phylobrain/bioinfo/sc_phylobrain/ggcorrconfplot.R")
```

```{r}
#Filtered Seurat
species1<-mouse_day1_orth
species2<-chicken_day1_orth
#Cell types order
species1_ct<-species1_ct
species2_ct<-species2_ct
```

```{r , echo=FALSE}
species1[['seurat_clusters']]<-species1[['species1_sct5_res2.8']]
species2[['seurat_clusters']]<-species2[['species2_2']]
Idents(species2)<-"seurat_clusters"
Idents(species1) <- "seurat_clusters"
```

#Regular Label Transfer

#Markers

Although this would have alredy been done in the past, it is easier to get it again than save all gene markers. Or maybe not, it could be removed then. 

```{r, echo=F}
DefaultAssay(species1)<-"RNA"
DefaultAssay(species2)<-"RNA"
species1.markers<-FindAllMarkers(species1_orth, only.pos = TRUE, min.pct = 0.25, logfc.threshold = 0.25)
species2.markers<-FindAllMarkers(species2_orth, only.pos = TRUE, min.pct = 0.25, logfc.threshold = 0.25)
```

#Top gene markers

Select the most differentially expressed genes, the top 400 of each cluster ordered by |avg_log2FC| and filtered by p_val_adj < 10^(-9). This was set based on Colquitt, but it could be changed. 

```{r, echo=F}
species1.markers_top<-species1.markers %>% group_by(cluster) %>% filter(p_val_adj < 10^(-9)) %>% arrange(abs(avg_log2FC)) %>% top_n(n=400) %>% dplyr::select(gene)
table(species1.markers_top$cluster)

species2.markers_top<-species2.markers %>% group_by(cluster) %>% filter(p_val_adj < 10^(-9)) %>% arrange(abs(avg_log2FC)) %>% top_n(n=400) %>% dplyr::select(gene) 
table(species2.markers_top$cluster)
```

The gene specificity index (GSI) is obtained for each species independently:

#Top Markers Subset - species2


```{r, echo=F}
species2_top<-species2_orth
species2_top
species2_top[["SCT"]]@counts<-species2_top[["SCT"]]@counts[species2.markers_top$gene,]
species2_top[["SCT"]]@data<-species2_top[["SCT"]]@data[species2.markers_top$gene,]
species2_top[["SCT"]]@meta.features[species2.markers_top$gene,'cluster']<-species2.markers_top$cluster
species2_top
```
#g coefficient - species2

```{r, echo=F}
Idents(species2_top)<-"seurat_clusters"
species2_cluster.averages <- AverageExpression(species2_top, assays = "SCT")
Idents(species2_top)<-"SCT_snn_res.0"
species2_averages <- AverageExpression(species2_top, assays = "SCT")
species2_g<-species2_cluster.averages$SCT/species2_averages$SCT[,1]
head(species2_g)
```


#Top Markers Subset - species1


```{r, echo=F}
species1_top<-species1_orth
species1_top
species1_top[["SCT"]]@counts<-species1_top[["SCT"]]@counts[species1.markers_top$gene,]
species1_top[["SCT"]]@data<-species1_top[["SCT"]]@data[species1.markers_top$gene,]
species1_top[["SCT"]]@meta.features[species1.markers_top$gene,'cluster']<-species1.markers_top$cluster
species1_top
```

#g coefficient - species1

```{r, echo=F}
Idents(species1_top)<-"seurat_clusters"
species1_cluster.averages <- AverageExpression(species1_top, assays = "SCT")
Idents(species1_top)<-"SCT_snn_res.0"
species1_averages <- AverageExpression(species1_top, assays = "SCT")
species1_g<-species1_cluster.averages$SCT/species1_averages$SCT[,1]
head(species1_g)
```

#GSI Correlation

Once the GSIs for each species and genes is generated. We compared the tendencies of these genes across the two species.


```{r, echo=F}
species2_gf<-as.matrix(species2_g[rownames(species2_g) %in% rownames(species1_g),])
species2_gf<-species2_gf[,species2_ct]
#To avoid possible common names of celltypes
colnames(species2_gf)<-lapply(colnames(species2_gf), function(x) paste("S1", x, sep="-"))
species1_gf<-as.matrix(species1_g[rownames(species1_g) %in% rownames(species2_g),])
species1_gf<-species1_gf[,species1_ct]
colnames(species1_gf)<-lapply(colnames(species1_gf), function(x) paste("S2", x, sep="-"))
table(rownames(species1_gf) %in% rownames(species2_gf))
#Step7: Merge Expression Tables
geTable = merge(species2_gf,species1_gf, by='row.names', all=F)
rownames(geTable) = geTable$Row.names
geTable = geTable[,2:ncol(geTable)]
gsi.cor<-cor(geTable, method = "spearman")
gsi.cor
```

```{r, echo=F}
#8b:  Shuffle data
shuffled.cor.list = list()
gsi   <- txtProgressBar(1, 100, style=3)
nPermutations <- 100
for (i in 1:nPermutations){
  shuffled = apply(geTable[,1:ncol(species2_gf)],1,sample)
  shuffled2 = apply(geTable[,(ncol(species2_gf)+1):ncol(geTable)],1,sample)
  shuffled = cbind(t(shuffled),t(shuffled2))
  shuffled.cor = cor(shuffled,method="spearman")
  shuffled.cor.list[[i]] = shuffled.cor
  rm(list=c('shuffled','shuffled2','shuffled.cor'))
  if ((i %% 100) ==0){
    setTxtProgressBar(gsi, (i*100)/nPermutations)
  }
}

p.value.table = matrix(ncol=ncol(geTable), nrow = ncol(geTable))
rownames(p.value.table) = colnames(geTable)
colnames(p.value.table) = colnames(geTable)

shuffled.mean.table = matrix(ncol=ncol(geTable), nrow = ncol(geTable))
rownames(shuffled.mean.table) = colnames(geTable)
colnames(shuffled.mean.table) = colnames(geTable)

a = combn(1:ncol(geTable),2)
for (i in 1:ncol(a)){
  cor.scores = sapply(shuffled.cor.list,"[",a[1,i],a[2,i])
  shuffled.mean.table[a[1,i],a[2,i]] = mean(cor.scores)
  shuffled.mean.table[a[2,i],a[1,i]] = mean(cor.scores)
  p.value = mean(abs(cor.scores)>=abs(gsi.cor[a[1,i],a[2,i]]))
  p.value.table[a[1,i],a[2,i]] = p.value
  p.value.table[a[2,i],a[1,i]] = p.value
  rm(list=c('cor.scores','p.value'))
  setTxtProgressBar(gsi, (i*100)/ncol(a))
}
p.value.table[1:ncol(species2_gf), (ncol(species2_gf)+1):ncol(gsi.cor)]
neg.log10.p = -log10(p.value.table)
```


```{r}
#step9 "Overlap in Markers"
#for all pairs of cell-types genspecies2te list of genes that are at least 1.5x avg in both cells

#from above a = combn(1:ncol(geTable),2)
marker.overlap.list = list()
for (i in 1:ncol(a)){
  datasubset = cbind(geTable[,a[1,i]],geTable[,a[2,i]])
  markers = rownames(geTable[datasubset[,1]>1.5 & datasubset[,2]>1.5,])
  marker.overlap.list[[i]] = markers
  names(marker.overlap.list)[i] = paste(colnames(geTable)[a[1,i]], colnames(geTable)[a[2,i]],sep='_')
  rm(list=c('datasubset','markers'))
}
```

```{r}
#We remove the labels SP1/SP2 to make the plot with the original cell type names. The number should be modified dependently of the label le
colnames(gsi.cor)<-str_sub(colnames(gsi.cor), start = 4)
rownames(gsi.cor)<-str_sub(rownames(gsi.cor), start = 4)
```

The plot by default can be modified to obtain a more informative result.
We have generated two versions depending on the color corresponding to max and min expression.
The repeated chunk below each corrplot function is to generate a pdf out of it. 


This plot considers positive as blue and negative as red. It is the default option of corrplot.

```{r, fig.height=7, fig.width=10, echo=F}
#plot it 
#blue, max. red, min.
p_inv<-1-p.value.table
corrplot::corrplot(gsi.cor[(ncol(species2_gf)+1):ncol(gsi.cor),1:ncol(species2_gf)], order="original", tl.pos="lt", method="color", tl.col="black", is.corr=F,tl.cex=0.7, sig.level=(0.95),insig="pch", pch=5, p.mat=p_inv[(ncol(species2_gf)+1):ncol(gsi.cor),1:ncol(species2_gf)], pch.cex=0.15,pch.col="black",mar=c(3,1,5,1),cl.align.text="l")
pdf("sp1vs2_cor.pdf")
corrplot::corrplot(gsi.cor[(ncol(species2_gf)+1):ncol(gsi.cor),1:ncol(species2_gf)], order="original", tl.pos="lt", method="color", tl.col="black", is.corr=F,tl.cex=0.7, sig.level=(0.95),insig="pch", pch=5, p.mat=p_inv[(ncol(species2_gf)+1):ncol(gsi.cor),1:ncol(species2_gf)], pch.cex=0.15,pch.col="black",mar=c(3,1,5,1),cl.align.text="l")
dev.off()
```

This plot considers high as red, and low as blue, as a heatmap. 


```{r, fig.height=7, fig.width=10, echo=F}
#plot it 
#red, max. blue, min.
p_inv<-1-p.value.table
corrplot::corrplot(gsi.cor[(ncol(species2_gf)+1):ncol(gsi.cor),1:ncol(species2_gf)], order="original", tl.pos="lt", method="color", tl.col="black", is.corr=F,tl.cex=0.7, sig.level=(0.95),insig="pch", pch=5, p.mat=p_inv[(ncol(species2_gf)+1):ncol(gsi.cor),1:ncol(species2_gf)], pch.cex=0.15,pch.col="black",mar=c(3,1,5,1),cl.align.text="l",col=colorRampPalette(rev(brewer.pal(n=10, name = "RdBu")))(200))

pdf("sp1vs2_cor_heat.pdf")
corrplot::corrplot(gsi.cor[(ncol(species2_gf)+1):ncol(gsi.cor),1:ncol(species2_gf)], order="original", tl.pos="lt", method="color", tl.col="black", is.corr=F,tl.cex=0.7, sig.level=(0.95),insig="pch", pch=5, p.mat=p_inv[(ncol(species2_gf)+1):ncol(gsi.cor),1:ncol(species2_gf)], pch.cex=0.15,pch.col="black",mar=c(3,1,5,1),cl.align.text="l",col=colorRampPalette(rev(brewer.pal(n=10, name = "RdBu")))(200))
dev.off()
```
