---
title: "Integration Function"
output: html_notebook
---

```{r}
library(dplyr)
library(Seurat)
library(SeuratObject)
library(patchwork)
library(Matrix)
library(glmGamPoi)
library(sctransform)
library(limma)
library(clustree)
library(pvclust)
library(Hmisc)
library(corrplot)
library(ggcorrplot)
library(stringr)
library(biomaRt)
source("/media/phylobrain/Windows/Users/rodri/Documents/R/win-library/4.1/corrplot/R/corrtry.R")
source("/home/phylobrain/bioinfo/sc_phylobrain/ggcorrconfplot.R")
```

```{r}
#Filtered Seurat
species1<-mouse_day1_orth
species2<-chicken_day1_orth
#if a preferred order of clusters is preferred
species1_ct<-species1_ct
species2_ct<-species2_ct
#if a palette is stablished for each species UMAP
colors_sp1<-colors_sp1
colors_sp2<-colors_sp2
```
#Select the same number of cells if needed

```{r , echo=FALSE}
species1[['seurat_clusters']]<-species1[['species1_sct5_res2.8']]
species2[['seurat_clusters']]<-species2[['species2_2']]
Idents(species2)<-"seurat_clusters"
Idents(species1) <- "seurat_clusters"
```

```{r}
colors<-list(colors_sp1, colors_sp2) 
order<-list(species1_ct, species2_ct) 
initials<-list("SP1","SP2") 
```

#Regular Integration

```{r}
DefaultAssay(species1)<-"RNA"
species1.list <- SplitObject(species1, split.by = "orig.ident")
species1.list <- lapply(X = species1.list, FUN = SCTransform, method = "glmGamPoi",  vars.to.regress = c("percent.mt", "G2M.Score", "S.Score"), variable.features.n = 2000)
DefaultAssay(species2)<-"RNA"
species2.list <- SplitObject(species2, split.by="orig.ident")
species2.list <- lapply(X = species2.list, FUN = SCTransform, method = "glmGamPoi",  vars.to.regress = c("S.Score", "G2M.Score","percent.mt"), variable.features.n = 2000)
integrated.list <- c(species1.list, species2.list)
features <- SelectIntegrationFeatures(object.list = integrated.list, nfeatures = 2000)
integrated.list <- PrepSCTIntegration(object.list = integrated.list, anchor.features = features)
integrated.list <- lapply(X = integrated.list, FUN = RunPCA, features = features)
anchors <- FindIntegrationAnchors(object.list = integrated.list, reduction = "rpca", dims = 1:50, normalization.method = "SCT", anchor.features = features) #It could be selected "cca" as reduction
integrated <- IntegrateData(anchorset = anchors, normalization.method = "SCT", dims = 1:50)
```

```{r, echo=FALSE}
integrated[["Species"]]<-"Species"
#Identify the orig.ident specific for each species and modify conveniently the Species meta.data
table(integrated@meta.data$orig.ident)
integrated[["Species"]][grep("^s", rownames(integrated@meta.data)),]<-"Species1"
integrated[["Species"]][grep("^X", rownames(integrated@meta.data)),]<-"Species2"
#Confirm
table(integrated[["Species"]])
#Generate the cell_types meta.data with the seurat_clusters info, seurat_clusters will dissapear once reclustering is done
integrated[["cell_types"]]<-"Animals"
integrated[["cell_types"]]<- integrated@meta.data$seurat_clusters
table(integrated[["cell_types"]])
```


# Integration

```{r , echo=FALSE}
integrated <- RunPCA(integrated, npcs = 50, verbose = FALSE)
integrated <- FindNeighbors(integrated, dims = 1:50)
#These resolutions are iterative, you must tune the parameter to obtain the representation of the cell types present in the original atlases.
integrated <- FindClusters(integrated, resolution = c(2,2.5,3))
integrated <- RunUMAP(integrated, dims = 1:50)
```

#Integrated Atlas

```{r}
DimPlot(integrated, group.by = "Species")
```

```{r}
DimPlot(integrated, group.by = "cell_types")
```

#Individual subsets of the integrated UMAP by cell_types

```{r}
n<-unique(integrated@meta.data$Species)
for(i in 1:length(n)){
  Idents(integrated)<-"Species"
  indiv<-subset(integrated, idents= n[i])
  DimPlot(indiv, group.by = "cell_types")
  Idents(indiv)<-"cell_types"
  levels(indiv)<- lapply(order[[i]], function(x) paste(initials[[i]], x, sep="-"))
  print(DimPlot(indiv, label = T)+NoLegend() # +scale_color_manual(values = unname(colors[[i]])))
}
#Save the plots as pdf
pdf("integrated_Sp1vsSp2_cell_types.pdf", width = 9, height = 6)
n<-unique(integrated@meta.data$Species)
for(i in 1:length(n)){
  Idents(integrated)<-"Species"
  indiv<-subset(integrated, idents= n[i])
  DimPlot(indiv, group.by = "cell_types")
  Idents(indiv)<-"cell_types"
  levels(indiv)<- lapply(order[[i]], function(x) paste(initials[[i]], x, sep="-"))
  print(DimPlot(indiv, label = T)+NoLegend() # +scale_color_manual(values = unname(colors[[i]])))
}
dev.off()
```

#Gene Expression

Visualization of marker genes for each cluster 

```{r}
genes_fig5d<-c("Pdgfra","Aldoc","Ascl1","Dlx1","Neurod2","Reln")
for(i in 1:length(genes_fig5d)){print(FeaturePlot(bs_e12p3_all, features = genes_fig5d[i]))}
```

#Reclusterized Cell Types

```{r}
DimPlot(integrated, group.by = "integrated_snn_res.2")
```

#Individual subsets of the integrated UMAP by integrated_snn_res.2

```{r}
n<-unique(integrated@meta.data$Species)
for(i in 1:length(n)){
  Idents(integrated)<-"Species"
  indiv<-subset(integrated, idents= n[i])
  DimPlot(indiv, group.by = "integrated_snn_res.2")
  Idents(indiv)<-"integrated_snn_res.2"
  levels(indiv)<- lapply(order[[i]], function(x) paste(initials[[i]], x, sep="-"))
  print(DimPlot(indiv, label = T)+NoLegend())
}
#Save the plots as pdf
pdf("integrated_Sp1vsSp2_newres.pdf", width = 9, height = 6)
n<-unique(integrated@meta.data$Species)
for(i in 1:length(n)){
  Idents(integrated)<-"Species"
  indiv<-subset(integrated, idents= n[i])
  DimPlot(indiv, group.by = "integrated_snn_res.2")
  Idents(indiv)<-"integrated_snn_res.2"
  levels(indiv)<- lapply(order[[i]], function(x) paste(initials[[i]], x, sep="-"))
  print(DimPlot(indiv, label = T)+NoLegend())
}
dev.off()
```

#ReNaming New Clusters

```{r}
Idents(integrated)<-"integrated_snn_res.2"
DimPlot(integrated, reduction = "umap", label = TRUE, pt.size = 0.5) + NoLegend()
integrated <- RenameIdents(integrated, "27"="","17"="", "30"="", "10"="", "2"="","25"="","19"="","22"="","18"="", "3"="","20"="", "8"="", "6"="","5"="","28"="","12"="","7"="", "21"="","11"="", "32"="","1"="","16"="")
integrated[["integrated"]]<-Idents(integrated)
p1<-DimPlot(integrated, reduction = "umap", label = TRUE, pt.size = 0.5) + NoLegend()
p1
```

```{r}
#Determine the desired order of clusters (integrated_ct)
integrated_ct<-c("","")
levels(integrated)<-integrated_ct
saveRDS(integrated, "integrated.rds")
```


# Species Proportion - BarPlot


```{r}
Idents(integrated)<-"integrated"
per<-table(integrated@meta.data$integrated, integrated@meta.data$Species)
percent<-data.frame(per[,1], per[,2])
colnames(percent)<-c("Species1","Species2")
percent2<-percent
for(i in 1:ncol(percent2)){
  percent2[,i]<-percent2[,i]/colSums(percent2)[i]
}
percent3<-percent2/rowSums(percent2)
percent3$Clusters<-rownames(percent3)
percent4<-percent3 %>% pivot_longer(cols = c(Species1, Species2), names_to = "Species", values_to = "Percent")
percent4$Round<-round(percent4$Percent*100)
percent4$Percent<-round(percent4$Percent, digits = 3)
#percent4$Clusters<-factor(percent4$Clusters, levels = sort(unique(percent4$Clusters)))
percent_int <- percent4
p1<-percent_int %>% ggplot(aes(x = Clusters, y = Percent, fill = factor(Species, levels = c("Species1","Species2"))))+
  geom_bar(stat = "identity") + geom_text(aes(label = paste(Round,"%"), y = Percent), position = position_stack(vjust = 0.5))+  coord_flip()+ labs(x = "Cluster", y = "Percentage",fill = "Species")+scale_fill_manual(values = c("steelblue3", "firebrick2"))
p1
```

Save the plot

```{r, fig.height=10, fig.width=7}
pdf("proportion_Sp1vsSp2.pdf", width = 7, height = 10)
percent_int %>% ggplot(aes(x = factor(Clusters, levels = rev(pb_ct)), y = Percent, fill = factor(Species, levels = c("Species1","Species2"))))+
  geom_bar(stat = "identity") + geom_text(aes(label = paste(Round,"%"), y = Percent), position = position_stack(vjust = 0.5))+  coord_flip()+ labs(x = "Cluster", y = "Percentage",fill = "Species")+scale_fill_manual(values = c("steelblue3", "firebrick2"))
dev.off()
``` 


# New Cell Types Dendrogram 

For creating this dendrogram, it is needed to run the barplot code.
The colors of the branches depend on the proportion of cells coming from the original species.
Blue dominated is Mouse and Red dominated is Chicken, Purple is Mixed. 


```{r}
integrated.data<-AverageExpression(integrated, assays="SCT", group.by = "integrated")$SCT
integrated.cor<-cor(integrated.data, integrated.data, method = "spearman")
integrated.cor
```


```{r, fig.height=4, fig.width=10, echo=F}
library(dendextend)
integrated.cor.dist<-dist(integrated.cor)
dend<-hclust(integrated.cor.dist, method = "ward.D2")
order<-integrated_ct
colors<-dend$order #it has no meaning, it is just to create a vector with the proper length
for(i in 1:length(dend$order)){
cluster<-percent4[percent4$Clusters == order[i],]
blue<-cluster[cluster$Species == "Mouse","Percent"]/100
red<-cluster[cluster$Species == "Chicken","Percent"]/100
if(blue>0.5){
colors[i]<-rgb(red*2, blue=0, 1) } else{
  colors[i]<-rgb(1, blue=0, blue*2)
}
}
col2rgb(colors, alpha = FALSE)
colors
```

```{r}
dend %>% as.dendrogram()  %>% set("branches_col", value="grey") %>% set("branches_k_col", value=colors, k=10) %>% set("leaves_pch", 17) %>% set("leaves_cex", 2) %>% set("leaves_col", value=colors, k=9) %>% set("branches_lwd", 4) %>% plot()
pdf("dendrogram_Sp1vsSp2.pdf", height = 4, width = 10)
dend %>% as.dendrogram()  %>% set("branches_col", value="grey") %>% set("branches_k_col", value=colors, k=10) %>% set("leaves_pch", 17) %>% set("leaves_cex", 2) %>% set("leaves_col", value=colors, k=9) %>% set("branches_lwd", 4) %>% plot()
dev.off()
```
