---
title: "Integration Function"
output: html_notebook
---

```{r}
library(dplyr)
library(Seurat)
library(SeuratObject)
library(patchwork)
library(Matrix)
library(glmGamPoi)
library(sctransform)
library(limma)
library(clustree)
library(pvclust)
library(Hmisc)
library(corrplot)
library(ggcorrplot)
library(stringr)
library(biomaRt)
source("/media/phylobrain/Windows/Users/rodri/Documents/R/win-library/4.1/corrplot/R/corrtry.R")
source("/home/phylobrain/bioinfo/sc_phylobrain/ggcorrconfplot.R")
```

```{r}
#Filtered Seurat
integrated<-readRDS("integrated.rds")
#if a preferred order of clusters is preferred
species1_ct<-species1_ct
species2_ct<-species2_ct
#if a palette is stablished for each species UMAP
colors_sp1<-colors_sp1
colors_sp2<-colors_sp2
```

```{r , echo=FALSE}
species1[['seurat_clusters']]<-species1[['species1_sct5_res2.8']]
species2[['seurat_clusters']]<-species2[['species2_2']]
Idents(species2)<-"seurat_clusters"
Idents(species1) <- "seurat_clusters"
```

```{r}
colors<-list(colors_sp1, colors_sp2) 
order<-list(species1_ct, species2_ct) 
initials<-list("SP1","SP2") 
```

# Markers Selection

This tool was developed based on the SCT data of the integrated object. 

```{r, echo=F}
DefaultAssay(integrated)<-"SCT"
integrated.list<-SplitObject(integrated, split.by = "Species")
species1.makers<-FindAllMarkers(integrated.list$species1, only.pos = TRUE, min.pct = 0.25, logfc.threshold = 0.25)
species2.makers<-FindAllMarkers(integrated.list$species2, only.pos = TRUE, min.pct = 0.25, logfc.threshold = 0.25)
integrated.makers <-rbind(species1.makers,species2.makers)
integrated.makers_mod <-integrated.makers %>% group_by(cluster) %>% filter(p_val_adj < 10^(-9)) %>% top_n(n=10, wt = avg_log2FC) %>% dplyr::select(gene)
```

However, this data can be obtained equivalently from SCT/Integrated of each species.

#Modules generation

```{r}
length<-length(unique(integrated.makers_mod$cluster))
for(i in 1:length){
  celltype<-as.character(unique(integrated.makers_mod$cluster)[i])
  cell.type<-str_replace(celltype, "-", ".")
  features<- integrated.makers_mod %>% filter(cluster == celltype) 
  integrated<-PercentageFeatureSet(integrated, assay= "SCT", features = features$gene, col.name = cell.type)
  mean<-mean(integrated@meta.data[WhichCells(integrated, idents = celltype),cell.type])
  integrated[[cell.type]]<-integrated@meta.data[[cell.type]]/mean
}
#Visualize modules stored in metadata (column number can vary)
integrated@meta.data[,c(52:84)]
```

To visualize, we used DotPlat, a modified version of Seurat::DotPlot, that does not modify the counts stored. 

```{r}
DotPlat(integrated,  features = colnames(integrated@meta.data[,c(52:84)]), scale=F)+ RotatedAxis()+ scale_color_gradientn(colours = c("grey90", "gold","dark green"), limits=c(0,1))
pdf("DotPlot_integrated.pdf", width = 10, height = 10)
DotPlat(integrated,  features = colnames(integrated@meta.data[,c(52:84)]), scale=F)+ RotatedAxis()+ scale_color_gradientn(colours = c("grey90", "gold","dark green"), limits=c(0,1))
dev.off()
```

#If Transcription Factor Only Modules

You have to recover the Ensembl objects: TFdata_species1 (we assume that these datasets are conserved, but it can be modified to be specific)

```{r}
#we recover the original integrated.makers
integrated.makers_TF<-merge(y=TFdata_species1, x=integrated.makers, by.y= "external_gene_name", by.x= "gene")
#We filter again by the top4
integrated.makers_mod <-integrated.makers_TF %>% group_by(cluster) %>% filter(p_val_adj < 10^(-9)) %>% top_n(n=4, wt = avg_log2FC) %>% dplyr::select(gene) %>% arrange(cluster)
```

```{r}
length<-length(unique(integrated.makers_mod$cluster))
for(i in 1:length){
  celltype<-as.character(unique(integrated.makers_mod$cluster)[i])
  cell.type<-str_replace(celltype, "-", ".")
  features<- integrated.makers_mod %>% filter(cluster == celltype) 
  integrated<-PercentageFeatureSet(integrated, assay= "SCT", features = features$gene, col.name = cell.type)
  mean<-mean(integrated@meta.data[WhichCells(integrated, idents = celltype),cell.type])
  integrated[[cell.type]]<-integrated@meta.data[[cell.type]]/mean
}
#Visualize modules stored in metadata (column number can vary)
integrated@meta.data[,c(52:84)]
```

To visualize, we used a modified version of Seurat::DotPlot, that does not modify the counts stored. 

```{r}
DotPlat(integrated,  features = colnames(integrated@meta.data[,c(52:84)]), scale=F)+ RotatedAxis()+ scale_color_gradientn(colours = c("grey90", "gold","dark green"), limits=c(0,1))
pdf("DotPlot_integrated_TF.pdf", width = 10, height = 10)
DotPlat(integrated,  features = colnames(integrated@meta.data[,c(52:84)]), scale=F)+ RotatedAxis()+ scale_color_gradientn(colours = c("grey90", "gold","dark green"), limits=c(0,1))
dev.off()
```
